---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=TRUE)
require(dplyr)
require(ggplot2)
```

## Loading and preprocessing the data

1. Load the data
```{r}
act <- read.csv(unzip('activity.zip', 'activity.csv'))
```

## What is mean total number of steps taken per day?
1. Histogram of the total number of steps taken each day
```{r}
total <- aggregate(steps ~ date, act, sum)
hist(total$steps, main='Histogram of Total steps taken each day')
```

2. Calculate and report mean and median total number of steps taken per day
```{r}
mean(total$steps)
median(total$steps)
```

## What is the average daily activity pattern?
1. Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
plot(act$interval, act$steps,pch=20,col="lightgray")
abline(h=mean(act$steps,na.rm=TRUE), col="black")
```

2. The 5-minute interval that contains the maximum number of steps is
```{r}
max.step <- max(act$steps, na.rm=TRUE)
act.narm <- act[complete.cases(act$steps),]
act.narm[act.narm$steps==max.step,]$interval
```
## Imputing missing values
1. The total number of missing values in the dataset is
```{r}
sum(!complete.cases(act))
```
2. Filling in all the missing values in the dataset.
* The missing number in the ``steps`` column
```{r}
sum(is.na(act$steps))
```
* Since the number of NA in ``steps`` column is equal to the number of missing values, we only need to filling in the missing values in the ``steps`` column.

3. Create a new data set with missing steps filling in with mean of 5-minute interval.
```{r}
mean.step = mean(act$steps,na.rm=TRUE)
full <- act %>%
          mutate( steps = ifelse(is.na(steps), mean.step, steps) )
# head(full)
# sum(!complete.cases(full))
# plot(full$interval, full$steps,pch=20,col="lightgray")
# abline(h=mean(full$steps,na.rm=TRUE), col="black")
```

4. Histogram of the total number of steps taken each day
```{r}
total.full <- aggregate(steps ~ date, full, sum)
hist(total.full$steps, main='Histogram of Total steps taken each day')
```

* Calculate and report mean and median total number of steps taken per day
```{r}
mean(total.full$steps)
median(total.full$steps)
```

* The mean is the same.
* The median is different.

The imputing missing data does not have any impact on the mean but have very slightly effect on the median.

## Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor ``daytype`` with two levels ``weekday`` and ``weekend``
```{r}
act <- act %>%
  mutate( daytype = ifelse( 
      weekdays(as.Date(date),TRUE) %in% c("Sat","Sun"),
      'weekend',
      'weekday'
    )
  )
```
2. Plot weekday vs weekend.
```{r}
ggplot(act, aes(interval, steps)) +
  geom_line() + facet_grid(rows=vars(daytype))